---
title: "P8105 HW 1 EP3045"
output: html_document
date: "2022-09-23"
---

#Problem 1

```{r include = FALSE}
library(tidyverse)
library(ggplot2)

```

```{r load penguin dataset}
library(palmerpenguins)
data("penguins", package = "palmerpenguins")
```

The names: 
`r names(penguins)`.

The unique *species*: 
`r unique(penguins$species)`.

The unique *islands*: 
`r unique(penguins$island)`.

The unique *years*: 
`r unique(penguins$year, na.rm = TRUE)`.

The *sex* of the penguins: 
`r unique(na.omit(penguins$sex))`.

The mean for *bill length* is `r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`mm, and the standard deviation is `r round(sd(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`.

The mean for *bill depth* is `r round(mean(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`mm, and the standard deviation is `r round(sd(pull(penguins, bill_depth_mm), na.rm = TRUE), digits=2)`.

The mean for *body mass* is `r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`mm, and the standard deviation is `r round(sd(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`.

The mean for *flipper length* is `r round(mean(pull(penguins, flipper_length_mm),na.rm = TRUE))`mm, and the standard deviation is `r round(sd(pull(penguins, flipper_length_mm), na.rm = TRUE), digits=2)`.

There are `r nrow(penguins)` *rows* and `r ncol(penguins)` *columns* in this dataset.

#### Creating a scatter plot

```{r create scatter plot, message = FALSE, warning = FALSE}
ggplot(penguins,aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + ggtitle("Flipper Length vs Bill Length")

#This is the scatter plot for Flipper versus Bill Length.
``` 
#### Saving the scatter plot

```{r save scatter plot, message = FALSE, warning = FALSE}
ggsave("penguin_scatter.pdf", plot = last_plot(), height = 4, width = 6)
```

This saves the scatter plot as a pdf file names penguin_scatter_plot.


#### Creating a data frame that consists of numeric, logical, character, and factor vectors.

```{r creating data frame}
set.seed(420)
data_table <- tibble(
  my_variable = rnorm(10),
  logical_vec = my_variable > 0,
  ten_char_vector = c("s", "t", "r", "a", "w", "b", "e", "r", "r", "y"),
  three_levels = factor(c("small", "medium", "large", "large", "medium", "small", "medium", "small", "large", "medium"))
)
data_table
```

#### Attempting to take the means of the variables

```{r finding mean of the variables}
mean(pull(data_table, my_variable))
mean(pull(data_table, logical_vec))
mean(pull(data_table, ten_char_vector))
mean(pull(data_table, three_levels))
```

The variables "my_variable" and "logical_vec" are numeric and logical vectors. Because of this, we are able to calculate their means. The variables "ten_char_vector" and "three_levels" are character and factor vectors. Because of this, we cannot calculate their means. 

#### Converting logicial, character, and factor variables using as.numeric without showing the outputs

```{r converting to variables to numeric, message = FALSE, warning = FALSE}
converted_num1 = as.numeric(pull(data_table, logical_vec))
converted_num2 = as.numeric(pull(data_table, ten_char_vector))
converted_num3 = as.numeric(pull(data_table, three_levels))
```

#### Finding the means of the converted variables

```{r finding converted mean}
mean(converted_num1)
mean(converted_num2)
mean(converted_num3)
```

We are able to convert logical and factor vectors into numeric values, but we are not able to convert character values into numeric values. The logical vector can be converted into a numeric value by assigning 0 to FALSE and 1 to TRUE. With this knowledge, I was able to convert the variables "logical_vec", "ten_char_vector", and "three_levels" into numeric values. Because "ten_char_vector" is a character vector, I cannot convert it into numeric variables. 